Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-11-11 paper/manuscript.tex
updating figures and manuscript

\begin{abstract}
Alchemical free energy calculations can be a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of modified potential energy functions to represent \emph{alchemical} intermediate states that cannot exist in chemistry; by analyzing simulation data collected from a series of bridging alchemical thermodynamic states, transfer free energies (or differences in transfer free energies) can be computed with orders of magnitude less simulation time than observing the process spontaneously. 
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen forcefield, and that appropriate corrections are included to permit direct comparison with experimental data.
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}

\end{frontmatter}
https://www.overleaf.com/5136551543vgjqbkzpmqkp

@@

%%%%%%%%%%%%%%%%%%%%
%  Introduction    %
%%%%%%%%%%%%%%%%%%%%
\section{What are alchemical free energy methods?}
\label{sec:intro}
Alchemical free energy (FE) calculations allow the calculation of energy differences associated with changes from one state to another, such as ligand binding to protein from solvent~\cite{zwanzig1954hightemperature}. The calculations use non-physical intermediate states where atoms are changed from one to another or can be removed or inserted by decoupling or coupling with their environment, Figure 1. These so-called alchemical states permit converged calculations for large changes, such as the entire removal of a ligand from a binding site, known as an absolute FE calculation, or modifying a substituent between two analogous ligands in a binding site, a relative FE calculation. Two common approaches are free energy perturbation (FEP) and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, biosimspace}. This foundation work, combined with the methodological, technological and hardware improvements of the last 5-10 years have led to an explosion of interest.  The domain of applicability for these calculations now involves such varied applications as the computation of protein-ligand binding free energies (both relative~\cite{relative-binding-free-energies} and absolute~\cite{absolute-binding-free-energies}), ligand selectivities~\cite{selectivity}, partition and distribution coefficients between different liquid phases (such as octanol-water partition coefficients~\cite{octanol-water-partition}), affinity changes due to mutations in therapeutic targets~\cite{hauser2018predicting,aldeghi2018accurate}, and changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}. As the field of molecular simulation moves towards millisecond timescales accurate FE calculations on even more challenging problems will be in scope. In the meantime, today's user may find it hard to get started with these complex calculations whilst also keeping up with the fast pace of change. This best practice guide aims to serve as a go-to reference for current recommendations and tips for users of all experience.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1/Fig1_dummy.pdf}
    \caption{{\bf What is alchemistry?}
    \todo[inline, color=green!20]{JDC: What is the concept for what this figure is supposed to communicate?}
    }
    \label{fig:fig1}
\end{figure}

@@

\section{Theoretical background}
\label{sec:theory}
\todo[inline, color={green!20}]{Info box what is free energy}
\todo[inline, color={green!20}]{Info box what is a thermodynamic cycle}

$\left[
      \begin{tabular}{@{\quad}m{0.8\columnwidth}@{\quad}}
          \raggedright%
          \textbf{What is a thermodynamic cycle?} \par
            There's a package called \texttt{url} designed for typesetting web addresses.
            Write \mbox{\texttt{\string\usepackage\string{url\string}}} in your preamble; this will provide the
            command \mbox{\texttt{\string\url}}. This command takes an address for the argument and
            will print it out with typewriter font. Furthermore, it is able to handle
            special characters in addresses like underscores and percent signs. It even
            enables hyphenation in addresses, which is useful for websites with a very
            long name.%
      \end{tabular}
\right]$
\newline
%May need to look at this again: https://tex.stackexchange.com/questions/66820/how-to-create-highlight-boxes-in-latex
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int dx \, e^{-u(x; \lambda)} .
\label{equation:partition-function-definition}
\end{eqnarray}
Here, the inverse thermal energy $\beta \equiv (k_B T)^{-1}$ where $k_B$ is the Boltzmann constant and $T$ is the absolute temperature, and the \emph{reduced potential} $u(x; \lambda)$~\cite{shirts2008statistically} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(x;\lambda) &\equiv& \beta \left[ U(x;\lambda) + p \, V(x) + \sum_{i=1}^N \mu_i \, N_i(x) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity.

\begin{figure}
    \includegraphics[width=0.95\columnwidth]{paper/figures/fig2/Fig2.pdf}
    \caption{{\bf Thermodynamic cycle for computing the relative free energy of binding ($\Delta \Delta G$) between two related small molecules to a supramolecular host.}
    The relative binding free energy difference between two small molecules, $\Delta \Delta G_{\mathrm{bind}, A \rightarrow B} \equiv \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A}$---here benzyl alcohol (top) to toluene (bottom)---can be computed as a difference between two alchemical transformations, $\Delta G_\mathrm{bound} - \Delta G_\mathrm{solvated}$, where $\Delta G_\mathrm{bound}$ represents the free energy change of transforming $A \rightarrow B$ in complex and $\Delta G_\mathrm{solvated}$ the free energy change of transforming $A \rightarrow B$ in solvent.}
    \label{fig:fig1}
\end{figure}
\todo[inline, color=green!20]{This section needs to be expanded and give a good introduction to the theoretical background without going into too much detail.}

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations? -- Step 0}
\label{sec:step0}
\todo[inline, color={green!20}]{@everyone please review this section}
Step 0 is about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are current alchemical free energy calculations?}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. An overview of benchmark datasets and case studies can be found in section \ref{sec:benchmark} at the end of this document.
\todo[inline,color={green!20}]{JDC: We should probably provide a short list of key benchmark references here, broken down by force field, and keep this up to date. GT: would a table, at the end of the document be an easier way to do this, only requiring some brief comments in the text? ASJSM: I like the table idea. MK: Feel free to remove this todo item if you are happy with the table provided. I refrained from just listing lots of case studies, as I think we should highlight the recommended datasets which should be used for benchmarking. Nevertheless, I have cited an review article containing such a summary as well as a recent publication comparing FEP+ and Gromacs in the brief text below the table.}
However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}. Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis} ASJSM: Can we incorporate your analysis in this section as an example?}
We thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.
%
\subsection*{Is the system of study suitable for alchemical free energy calculations?}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.

@@

\subsection*{Is the expected accuracy of the computation good enough?}
Additionally, the requisite level of accuracy is important -- if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracies in the 1-2 kcal/mol range~\cite{mobley2012perspective}, but if the number of compounds to be
synthesized is very small, this accuracy may not be enough to provide
much value.

@@

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/fig2/Fig2.pdf}
    \caption{a) Bad alignment of congeneric series b) good alignment}
    \label{fig:fig1}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/fig6/Fig3.pdf}
    \caption{Absolute thermodynamic cycle}
    \label{fig:fig1}
\end{figure}

@@

\section{Step 3 -- Overview of available analysis techniques}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BA: uncertainty estimation section}
\begin{enumerate}

\item Detecting boundary between equilibrated and production regions (Chodera: \url{http://dx.doi.org/10.1021/acs.jctc.5b00784})
\item Decorrelating samples for analysis
\todo[inline]{JDC: write section}
\begin{itemize}
\item Subsample different lambdas based on correlation times
\item Ensure all simulations at least 50x correlation time
\end{itemize}
\todo[inline]{MS: write section}

\todo[inline]{MS/JDC: write section}
\item Free energy differences between two different states differing in the energy function are directly related to the
ratio of probabilities of those states.AS can be noted, the partition functions in Eq.~\ref{equation:dimensionless-free-energy-difference} are simply the total accumulated probabilites for all possible configurations of the system. Virtually all of the ways to estimate this free energy are based in converting this ratio of integrals to something that can be measures in one (or several) simulations.

\begin{itemize}
    \item \textbf{The Zwanzig relationship} Perhaps the simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{orig.TPT}, also called EXP or simply free energy perturbation, thought that term later grew to encompass all ways of calculating free energy differences.
The free energy between two states defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - U_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta U(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
In words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be to switch to potential energy function $U_1(\vec{q})$, and exponentially average them to get the free energy.  Here, the exponential average consists in multiplying by -1, exponentiating, averaging, taking the logarithm, and multiplying by -1 again.  This almost requires no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies.  The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change some conditions~\cite{xx} 
Although the Zwanzig equation is formally correct (as long as the two states considered sample the same phase space volume), it has some very important numerical issues that mean that it is usually when performing a standard free energy calculation for a small molecule.~\cite{shirts.comparison,LuND:Appmcf}  One can show that if the standard deviation of the difference $\Delta U(\vec{q}) = U_1(vec{q})-U_2(\vec{q})$ over all sampled $\vec{q}$ is large (which in this case, means only several times $k_BT)$, then very few samples contribute to the average, and the answer will be both biased and extremely noisy~\cite{xx}.  There are a number of applications where it is useful, but those usually involve investigating the dependence of the free energy on small changes in parameters, not changing molecules or even single atoms in molecules~\cite{xx}

\item Bennett Acceptance Ratio

>>>>

\begin{abstract}
Alchemical free energy calculations can be a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of bridging potential energy functions representing \emph{alchemical} intermediate states that cannot exist in chemistry; by analyzing simulation data collected from these bridging alchemical thermodynamic states, transfer free energies (or differences in transfer free energies) can be computed with orders of magnitude less simulation time than simulating the transfer process directly. 
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen force field, and that appropriate corrections are included to permit direct comparison with experimental data.
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}

\end{frontmatter}

@@

%%%%%%%%%%%%%%%%%%%%
%  Introduction    %
%%%%%%%%%%%%%%%%%%%%

\todo[inline]{}The current Overleaf document is \url{https://www.overleaf.com/5136551543vgjqbkzpmqkp}}

\section{Abbreviations}
\label{sec:abbreviations}

\todo[inline]{JDC: Should we add an Abbreviations section? Maybe it can be in a shaded box?}

\section{What are alchemical free energy methods?}
\label{sec:intro}
Alchemical free energy (FE) calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor or the transfer of a small molecule from an aqeuous to apolar phase~\cite{zwanzig1954hightemperature}. 
The calculations use non-physical intermediate states where atoms are changed from one to another or can be removed or inserted by decoupling or coupling with their environment, Figure~\ref{fig:}. These so-called alchemical states permit converged calculations for large changes, such as the entire removal of a ligand from a binding site, known as an absolute FE calculation, or modifying a substituent between two analogous ligands in a binding site, a relative FE calculation. Two common approaches are free energy perturbation (FEP) and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, biosimspace}. This foundation work, combined with the methodological, technological and hardware improvements of the last 5-10 years have led to an explosion of interest.  The domain of applicability for these calculations now involves such varied applications as the computation of protein-ligand binding free energies (both relative~\cite{relative-binding-free-energies} and absolute~\cite{absolute-binding-free-energies}), ligand selectivities~\cite{selectivity}, partition and distribution coefficients between different liquid phases (such as octanol-water partition coefficients~\cite{octanol-water-partition}), affinity changes due to mutations in therapeutic targets~\cite{hauser2018predicting,aldeghi2018accurate}, and changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}. As the field of molecular simulation moves towards millisecond timescales accurate FE calculations on even more challenging problems will be in scope. In the meantime, today's user may find it hard to get started with these complex calculations whilst also keeping up with the fast pace of change. This best practice guide aims to serve as a go-to reference for current recommendations and tips for users of all experience.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1/Fig1_dummy.pdf}
    \caption{{\bf What is alchemistry?}
    \todo[inline, color=green!20]{JDC: What is the concept for what this figure is supposed to communicate?}
    }
    \label{fig:illustrative-alchemical-transformation}
\end{figure}

@@

\section{Statistical mechanics can tell us why alchemical free energy calculations work}
\label{sec:theory}
Why would want to run an alchemical free energy calculation and why do they work? Suppose you want to compute the binding affinity, or free energy of binding, of a ligand $L$ to a protein $P$, given by:
\begin{euqation}
P+L\leftrightharpoons PL
\end{euqation}
The equilibrium rate of binding is given by the ratio of concentrations of product $[PL]$ and reactants $[P]$, $[L]$:
\begin{equation}
 K_b^{\circ} = c^{\circ}\frac{[[PL]}{[L][P]}
\end{equation}
In order to define a reference state the, standard state is usually defined as $1 \mathrm{mol}/\mathrm{L}$ and $c^\circ$ is the standard state concentration. From the rate of binding a binding free energy is given by:
\begin{equation}
    \Delta G_{\mathrm{bind}} = -RT\ln K_b^{\circ}
\end{equation}
One way to estimate the binding rate is by computing the probability of finding the molecular system in the bound state of the unbound state, making use of partition function ratios, i.e.

\begin{equation}
    \Delta G_{\mathrm{bind}} = -RT\ln\frac{\int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}{\int_\mathrm{unbound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}
\end{equation}
Or more generally in a simplified form the free energy difference between the bound state is:
\begin{equation}
\Delta f \equiv f(\mathrm{bound}) - f(\mathrm{unbound}) = -\ln\frac{Z(\mathrm{unbound})}{Z({\mathrm{bound}})},
\end{equation}
where $Z = \int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}$ represents the configurational integral. 
How can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation. This is where alchemical free simulations come in. We can write down a thermodynamic cycle as seen in ~\ref{fig:illustrative-alchemical-transformation}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int dx \, e^{-u(x; \lambda)} .
\label{equation:partition-function-definition}
\end{eqnarray}
Here, the inverse thermal energy $\beta \equiv (k_B T)^{-1}$ where $k_B$ is the Boltzmann constant and $T$ is the absolute temperature, and the \emph{reduced potential} $u(x; \lambda)$~\cite{shirts2008statistically} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(x;\lambda) &\equiv& \beta \left[ U(x;\lambda) + p \, V(x) + \sum_{i=1}^N \mu_i \, N_i(x) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity.

\begin{figure}
    \includegraphics[width=0.95\columnwidth]{paper/figures/fig2/Fig2.pdf}
    \caption{{\bf Thermodynamic cycle for computing the relative free energy of binding ($\Delta \Delta G$) between two related small molecules to a supramolecular host.}
    The relative binding free energy difference between two small molecules, $\Delta \Delta G_{\mathrm{bind}, A \rightarrow B} \equiv \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A}$---here benzyl alcohol (top) to toluene (bottom)---can be computed as a difference between two alchemical transformations, $\Delta G_\mathrm{bound} - \Delta G_\mathrm{solvated}$, where $\Delta G_\mathrm{bound}$ represents the free energy change of transforming $A \rightarrow B$ in complex and $\Delta G_\mathrm{solvated}$ the free energy change of transforming $A \rightarrow B$ in solvent.}
    \label{fig:binding-thermodynamic-cycle}
\end{figure}
\todo[inline, color=green!20]{This section needs to be expanded and give a good introduction to the theoretical background without going into too much detail.}

%%%%%%%%%%%%
% Step 0   %
%%%%%%%%%%%%
\section{What can be expected from alchemical simulations? -- Step 0}
\label{sec:step0}

\todo[inline, color={green!20}]{@everyone please review this section}

\todo[inline]{JDC: Do we need explicit "Step 0", "Step 1", or can we use some other scheme to clearly delineate the steps one must follow?}

Step 0 is about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are alchemical free energy calculations?}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. An overview of benchmark datasets and case studies can be found in section \ref{sec:benchmark} at the end of this document.
\todo[inline,color={green!20}]{JDC: We should probably provide a short list of key benchmark references here, broken down by force field, and keep this up to date. GT: would a table, at the end of the document be an easier way to do this, only requiring some brief comments in the text? ASJSM: I like the table idea. MK: Feel free to remove this todo item if you are happy with the table provided. I refrained from just listing lots of case studies, as I think we should highlight the recommended datasets which should be used for benchmarking. Nevertheless, I have cited an review article containing such a summary as well as a recent publication comparing FEP+ and Gromacs in the brief text below the table.}
However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

It is worth noting that the accuracy of free energy calculations is highly variable across different protein targets, and likely across different ligand chemotypes as well. For instance, FEP+ with OPLS3 achieves an RMSE of 0.62 kcal/mol for a set of 21 compounds binding to JNK1 kinase, yet only an RMSE of 1.05 kcal/mol for a set of 34 compounds binding to P38$\alpha$ kinase~\cite{harder2016opls3}. Furthermore, perturbations for the same chemotype in different pockets of the BACE enzyme gave varied errors~\cite{keranen2017acylguanidine}
\todo[inline,color={green!20}]{JDC: It's important to be clear on what error is being reported: $\Delta G$ after shifting by a constant to minimize the RMSE, unshifted $\Delta G$, $\Delta \Delta G$ of computed edges, or $\Delta \Delta G$ of all edges. Note that computed edge $\Delta \Delta G$ RMSE significantly underestimates the error of all edge RMSE: \url{https://github.com/jchodera/jacs-dataset-analysis} ASJSM: Can we incorporate your analysis in this section as an example?}
We thus recommend that retrospective study for a particular target and a particular chemical series be performed to establish the relevant accuracy of free energy calculations for each particular application case.
%

\subsection*{How reproducible are alchemical free energy calculations?}

\todo[inline]{Discuss implications of the SAMPL6 SAMPLing paper~\cite{rizzi2019sampl6} on reproducibility between simulation packages and potential sensitivity to other calculation details.}

\subsection*{Is my problem suitable for alchemical free energy calculations?}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.

@@

\subsection*{Is the expected accuracy of the computation sufficient?}
Additionally, the requisite level of accuracy is important---if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracies in the 1--2 kcal/mol range~\cite{mobley2012perspective}, but if the number of compounds to be
synthesized is very small, this accuracy may not be enough to provide
much value.

@@

\todo[inline]{JDC: Should this be a Section or Subsection?}
\todo[inline]{JDC: This section needs some structure (subsections, paragraph) to break up the "wall of text".}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/fig2/Fig2.pdf}
    \caption{a) Bad alignment of congeneric series b) good alignment}
    \label{fig:not-sure-what-this-is}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/fig6/Fig3.pdf}
    \caption{Absolute thermodynamic cycle}
    \label{fig:absolute-thermodynamic-cycle}
\end{figure}

@@

\section{Step 3 -- Data analysis}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BA: uncertainty estimation section}

Once data has been collected from alchemical intermediates using a sampling algorithm described above, this data must be analyzed to produce an estimate of the transfer free energy (and its associated statistical uncertainty) for each edge of the thermodynamic cycle.
While a number of different estimators are available that will give consistent results under optimal circumstances, some approaches are recommended over others due to their robustness and ability to report on poor convergence.

\subsection*{Detecting the boundary between equilibrated and production regions}
\label{sec:automatic-equilibration-detection}

Much of the infrastructure for analyzing alchemical free energy calculations relies on the concept of asymptotically unbiased estimators, which produce unbiased estimates of the free energy when fed very long simulations~\cite{shirts2005comparison}.
In reality, free energy calculations are often initiated from highly atypical initial conditions (such as a protein-ligand geometry obtained from docking and subjected to a heuristic solvent placement scheme), and simulations are of a finite length dictated by available computational resources and computing demands.
As a result, these estimators can produce significantly biased estimates if fed the entirety of simulation data generated without further processing~\cite{chodera2016simple}.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{paper/figures2/automatic-equilibration-detection/argon-potential.pdf}
    \includegraphics[width=0.95\linewidth]{paper/figures2/automatic-equilibration-detection/argon-potential-reverse-cumulative-average.pdf}
    \caption{{\bf Automatic partitioning into equilibration and production regions.}
    \emph{Top:} The average (black line) standard deviation (shaded region) of the reduced potential $u^*$ over many independent replicate simulations started from the same initial conditions show a significant initial transient before relaxing to the true average potential energy.
    \emph{Middle:} A cumulative average (red) of the entire simulation data demonstrates a bias in the estimate of the average potential $<u>$ that persists far beyond the initial transient relaxation time, while averaging a production region that omits the initial transient (blue) eliminates this bias.
    \emph{Bottom:} A simple automated approach to determination of the equilibration boundary $t_0$ suggested in \cite{chodera2016simple} examines the number of effectively uncorrelated samples $N_\mathrm{eff}_{[t_0,T]} = (T-t_0) / g_{[t_0,T]}$ within the proposed production region $[t_0,T]$ by computing the statistical inefficiency $g_{[t_0,T]}$ for that simulation interval.
    $N_\mathrm{eff}_{[t_0,T]}$ has the general behavior of increasing linearly as $t_0$ moves backwards from $T$ until the initial transient relaxation is encountered, at which point it plunges.
    The optimal equilibration boundary $t_0$ is selected to maximize the number of uncorrelated samples.
    \emph{Figure adapted from \cite{chodera2016simple}.}
    \todo[inline]{JDC will update this figure with something similar for a real alchemical free energy calculation.}
    }
    \label{fig:automatic-equilibration-detection}
\end{figure}

To minimize this effect, an initial portion of the simulation is often discarded to \emph{equilibration}, with the idea of removing the most heavily biased initial portion of simulation data but retaining the unbiased \emph{production} region that represents a stationary Markov chain process sampling from the desired equilibrium target distribution.
Because the simulation time required for the atypical initial sampler state to relax toward equilibrium is a property of the specific system being simulated and the specific initial conditions selected, it is simplest to collect data for the whole process and use an automated algorithm to select how much data should be discarded to equilibration in a post-processing step.

A simple approach to automatically partitioning simulation data into equilibration and production regions is described in \cite{chodera2016simple} (illustrated in Figure~\ref{fig:automatic-equilibration-detection}).
Suppose we have a simulation of length $T$ consisting of correlated data.
Here, the goal of the post-processing step is to select the equilibration boundary $t_0 \in [0, T]$ so as to \emph{maximize} the number of effectively uncorrelated samples remaining in the production region $N_{[t_0,T]}$, which is defined as
\begin{eqnarray}
N_{[t_0,T]} &=& \frac{T - t_0}{g_{[t_0,T]}}
\end{eqnarray}
where $g_{[t_0,T]}$ is the \emph{statistical inefficiency} of a timeseries $a_t$, described in more detail below.
Conveniently, this procedure also produces the information necessary to decorrelate the simulation data for estimating the free energy differences, a requisite next step in analysis.
This approach is implemented within the MBAR~\cite{mbar} and alchemlyb~\cite{alchemlyb} packages, and is highly recommended for standard practice.

\paragraph{Computing the timeseries for equilibration detection}
Typically, the timeseries of note $a_t$ analyzed in automated equilibration detection is the negative logarithm of the probability density sampled by the MCMC algorithm (up to an irrelevant additive constant).
For simple independent simulations that sample $x_t \sim \pi(x ; \lambda)$, this is given by the reduced potential
\begin{eqnarray}
a_t &\equiv& - \ln \pi(x_t; \lambda) + c = u(x_t; \lambda) .
\end{eqnarray}
For a Hamiltonian replica exchange simulation where the super-replica state $(X_t, S_t) \sim \Pi(X_t, S_t)$, where $X \equiv \{x_1, \ldots, x_K\}$ are the positions and $S \equiv \{s_1, \ldots, s_K}$ is the alchemical state permutation vecftor, the negative logarithm of the super-replica reduced potential is used:
\begin{eqnarray}
a_t &\equiv& - \ln \Pi(X_t, S_t) + c = \sum_{k=1}^K u(x_{kt}; s_{kt})
\end{eqnarray}
Note that the use of the effective reduced potential is not guaranteed to pick up on all slow relaxation processes that may be coupled to the alchemical free energy, but the simplicity of its computation means it is generally appropriate for most cases.

\paragraph{Cautions in automating equilibration detection}
For simulations that are simply not long enough to contain a large number of samples from true equilibrium either because they are very short or contain slow processes, this procedure cannot completely remove the bias.
In those cases, this approach simply selects the final portion of the the simulation, which may be contained in a single substate of conformational space, which may itself lead to biased estimates. 
This situation can be detected if the equilibration boundary $t_0$ is a significant fraction of the total simulation length $T$, with a good rule of thumb being that $T \gtrsim 20 t_0$.
If this is not possible, advanced analysis techniques that assume only local equilibrium (rather than global equilibrium) such as the TRAM estimators~\cite{mey2014xtram,wu2016multiensemble,nuske2017markov} may be more appropriate.

\subsection*{Decorrelating samples for analysis}
\label{sec:decorrelating-samples}
